---
title: "Mapping_CrossNetwork_CUPNEx"
output: html_document
date: '2022-12-14'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Map 2: Tree Regeneration By Size Class
Map 5: Regeneration Composition
Map 6: Canopy Composition
Map 8: Priority Invasive Composition
Map 8b: Not done by NETN, but agreed that we'd all like a map of relative abundance of Native versus Exotic by plot.  Should be very straigtforward, two colors (one for native, one for exotic)

```{r}
library(tidyverse) 
library(dplyr)
library(httr)
library(rgdal) 
library(sf) 
library(RColorBrewer)
library(sp) 
library(rgeos)
library(leaflet)
library(leaflet.minicharts)
library(mapview)
library(lubridate)


setwd("C:/Users/kseeger/OneDrive - DOI/Desktop/GIT_RSTUDIO/SIP_NPS_CODE/Mapping/CUPN")

tree_basics <- read_csv("TreeBasics_20221013.csv")
seed_sap <- read_csv("SeedlingSapling_20221013.csv")
#seed_sap has become only CARL plots in this script

```

#Wrangling Seedling/Sapling Data for Map 2. Regeneration by Size Class
```{r}
seed_sap <- seed_sap %>%
dplyr::filter(grepl("CARL", Plot_Code)) %>%
dplyr::filter(Event_Type_Name != "QA/QC")

seed_sap_sum <- seed_sap %>%
  group_by(Plot_Code) %>%
  summarise(across(where(is.numeric), sum, na.rm = TRUE), .groups = 'drop')

seed_sum <- seed_sap_sum %>%
 dplyr::select(Plot_Code, starts_with("Seed"))


sap_sum <- seed_sap_sum %>%
 dplyr::select(Plot_Code, starts_with("Sap"))

```

#CUPN Plot Lat and Long
```{r}
cupn_plots <- read_csv("CUPN_PlotEvents_LatLong.csv") 

cupn_plots_select <- cupn_plots %>%
    dplyr::select(Plot_Code, Start_Date, Latitude, Longitude)

cupn_plots_select$Start_Date <- dmy(cupn_plots_select$Start_Date)

cupn_plots_select <- cupn_plots_select %>%
  group_by(Plot_Code) %>%
  slice(which.max(Start_Date))

cupn_carl <- cupn_plots_select %>%
    dplyr::filter(grepl("CARL", Plot_Code))

```

```{r}
#issue with barplot disliking the merge function and having NA values
#seed_sum <- merge(x = seed_sum, y = cupn_carl, all.y = T)

seed_sum <- left_join(seed_sum, cupn_carl, by = "Plot_Code")
seed_sum <- seed_sum %>%
  group_by(Plot_Code) %>%
dplyr::mutate(Sum = sum(across(Seedling_5_15_Tall:Seedling_50_137_Tall))) %>%
  ungroup()

sap_sum <- merge(x = sap_sum, y= cupn_carl, all.y = T)

sap_sum  <- sap_sum  %>%
  group_by(Plot_Code) %>%
dplyr::mutate(Sum = sum(across(Sapling_0_1_DBH:Sapling_5_10_DBH))) %>%
  ungroup()

```

#importing CARL Shapefile
```{r}
park_code <- "CARL"
  # KATE, ENTER 4-LETTER PARK CODE HERE
  
unitBoundaryURL_CARL <- paste0("https://services1.arcgis.com/fBc8EJBxQRMcHlei/ArcGIS/rest/services/IMD_Units_Generic_areas_of_analysis_(AOAs)_-_IMD_BND_ALL_UNITS_AOA_nad_py_view/FeatureServer/0/query?where=UNITCODE+%3D+%27", park_code, "%27&objectIds=&time=&geometry=&geometryType=esriGeometryEnvelope&inSR=&spatialRel=esriSpatialRelIntersects&distance=&units=esriSRUnit_Meter&relationParam=&outFields=*&returnGeometry=true&maxAllowableOffset=&geometryPrecision=&outSR=4326&gdbVersion=&returnDistinctValues=false&returnIdsOnly=false&returnCountOnly=false&returnExtentOnly=false&orderByFields=&groupByFieldsForStatistics=&outStatistics=&returnZ=false&returnM=false&resultOffset=&resultRecordCount=&f=geojson") # save output as WGS84

#paste(unitBoundaryURL)
```

```{r}
#checking error code below
imported_dat_carl <- tryCatch(readOGR(dsn = unitBoundaryURL_CARL, dropNULLGeometries = FALSE), error=function(e) print("Error retrieving data")) # return error message if problems still arise with downloading from web services

#shiny::req(class(imported_dat)=="SpatialPolygonsDataFrame")
imported_dat_carl <- sp::spTransform(imported_dat_carl, CRS("+proj=longlat +datum=WGS84")) # convert to WGS84
```

#Bar Plot Sum of Seedlings
#issues with bars being same height - issue arose after changing naming of variables
```{r}

#colorPalette = RColorBrewer::brewer.pal(colorblindFriendly, GnBu)
#changed some of the order etc
okabe_ito <- c( "#0072B2", "#D55E00", 
               "#CC79A7", "#000000",  '#CC6677', '#BBCC33', '#99DDFF', 
                '#332288', '#882255', '#FFAABB', "#E69F00", "#56B4E9", "#009E73", "#F0E442")

baseman_carl = leaflet(data = seed_sum) %>%
  addTiles() %>%
  addPolygons(data = imported_dat_carl, color = ~ "black")%>%
  addCircleMarkers(cupn_carl$Longitude,
                   cupn_carl$Latitude,  color = ~ "black", weight = 1, fillColor = ~"blue", fillOpacity = 1, radius = 3) %>%
  addLabelOnlyMarkers(cupn_carl$Longitude, 
                cupn_carl$Latitude, label =  cupn_carl$Plot_Code,
                      labelOptions = labelOptions(noHide = T, direction = 'top', textOnly = T, style=list('color'="#ffffff", 'fontSize'="10px")))


baseman_carl <- baseman_carl %>%
  addMinicharts(seed_sum$Longitude, 
                seed_sum$Latitude, 
    chartdata = dplyr::select(seed_sum, Seedling_5_15_Tall:Seedling_50_137_Tall), 
    col = okabe_ito,
    showLabels = F) %>%
    fitBounds(min(seed_sum$Longitude), min(seed_sum$Latitude), max(seed_sum$Longitude), max(seed_sum$Latitude))  %>%
  addLabelOnlyMarkers(seed_sum$Longitude, 
                seed_sum$Latitude, label =  seed_sum$Plot_Code,
                      labelOptions = labelOptions(noHide = T, direction = 'top', textOnly = T, style=list('color'="#ffffff", 'fontSize'="10px")))

baseman_carl


```

#Sum of Seedlings by size classes in each plot
```{r}
baseman_carl = leaflet(data = seed_sum) %>%
  addTiles() %>%
  addPolygons(data = imported_dat_carl, color = ~ "black")%>%
  addCircleMarkers(cupn_carl$Longitude,
                   cupn_carl$Latitude,  color = ~ "black", weight = 1, fillColor = ~"blue", fillOpacity = 1, radius = 4) %>%
  addLabelOnlyMarkers(cupn_carl$Longitude, 
                cupn_carl$Latitude, label =  cupn_carl$Plot_Code,
                      labelOptions = labelOptions(noHide = T, direction = 'top', textOnly = T, style=list('color'="#ffffff", 'fontSize'="10px")))


baseman_carl <- baseman_carl %>%
  addMinicharts(seed_sum$Longitude, 
                seed_sum$Latitude, 
    chartdata =dplyr::select(seed_sum, Seedling_5_15_Tall:Seedling_50_137_Tall), 
    type = "pie",
    width = sqrt(seed_sum$Sum) * 5,
    col = okabe_ito,
    showLabels = F) %>%
    fitBounds(min(seed_sum$Longitude), min(seed_sum$Latitude), max(seed_sum$Longitude), max(seed_sum$Latitude))  %>%
  addLabelOnlyMarkers(seed_sum$Longitude, 
                seed_sum$Latitude, label =  seed_sum$Plot_Code,
                      labelOptions = labelOptions(noHide = T, direction = 'top', textOnly = T, style=list('color'="#ffffff", 'fontSize'="10px")))

baseman_carl

## 'leaflet' objects (image above)

#mapshot(baseman_carl, file = "LeafletPieNum_CARL.png")

```

#Sum of Sapling by size class in each plot
```{r}
baseman_carl = leaflet(data = sap_sum) %>%
  addProviderTiles("Esri.WorldImagery", options = providerTileOptions(noWrap = TRUE), group="Satellite") %>%  
  addPolygons(data = imported_dat_carl, color = ~ "black") %>%
   addCircleMarkers(sap_sum$Longitude,
                   sap_sum$Latitude,  color = ~ "black", weight = 1, fillColor = ~"blue", fillOpacity = 1, radius = 4)

baseman_carl <- baseman_carl %>%
  addMinicharts(sap_sum$Longitude, 
                sap_sum$Latitude, 
    chartdata =dplyr::select(sap_sum, Sapling_0_1_DBH:Sapling_5_10_DBH), 
    type = "pie",
    width = sqrt(sap_sum$Sum) * 5,
    col = okabe_ito,
    showLabels = F) %>%
    fitBounds(min(sap_sum$Longitude), min(sap_sum$Latitude), max(sap_sum$Longitude), max(sap_sum$Latitude)) %>%
  addLabelOnlyMarkers(sap_sum$Longitude, 
                sap_sum$Latitude, label =  sap_sum$Plot_Code,
                      labelOptions = labelOptions(noHide = T, direction = 'top', textOnly = T, style=list('color'="#ffffff", 'fontSize'="10px")))

baseman_carl

```

#MAP 5 SEEDLING SAPLING REGENERATION COMPOSITION 
```{r}
#Finding the top 11 most occuring species
top_species <- seed_sap %>%
dplyr::mutate(Plant_Code = as.factor(Plant_Code)) %>%
  group_by(Plant_Code) %>%
  summarise(Observations = n()) %>%
  arrange(desc(Observations))  %>%
dplyr::mutate(Plant_Code = as.character(Plant_Code))

top_species <- as.data.frame(top_species[top_species$Observations >20,])

listspecies <- as.list(top_species[1])

seed_sap_top <- seed_sap %>%
dplyr::filter(Plant_Code %in% listspecies$Plant_Code)

seed_sap_top <- seed_sap_top %>%
  replace(is.na(.), 0)
```

#Counting occurance of top species 
```{r}
seed_sap_top <- seed_sap_top %>%
 dplyr::select(Plot_Code, Plant_Code, Seedling_5_15_Tall:Sapling_5_10_DBH) %>%
  group_by(Plot_Code, Plant_Code) %>%
  summarize(across(where(is.numeric), ~sum(.x, na.rm = TRUE))) %>%
  ungroup()

seed_sap_top <- seed_sap_top %>%
dplyr::mutate(Count = rowSums(across(where(is.numeric)))) %>%
 dplyr::select(Plot_Code, Plant_Code, Count)

seed_sap_top <- seed_sap_top %>%
  pivot_wider(values_from = Count, 
              names_from = Plant_Code)

seed_sap_top <- seed_sap_top %>%
dplyr::mutate(sum = rowSums(across(where(is.numeric)), na.rm= TRUE))
  

#seed_sap_top <- left_join(seed_sap_top, cupn_carl, by = "Plot_Code")

```
#Creating other variable for less prevalent species
```{r}
seed_sap_other <- seed_sap %>%
dplyr::filter(!Plant_Code %in% listspecies$Plant_Code)

seed_sap_other <- seed_sap_other %>%
  replace(is.na(.), 0)

seed_sap_other <- seed_sap_other %>%
dplyr::mutate(SapCount = rowSums(across(Seedling_15_30_Tall:Sapling_5_10_DBH)))

seed_sap_other <- seed_sap_other %>%
  dplyr::select(Plot_Code, SapCount) %>%
  group_by(Plot_Code) %>%
  summarize(Other = sum(SapCount))
  

```

#Calculating stem densities per hectare
```{r}
seed_sap_den <- seed_sap %>%
  dplyr::mutate(
    Seedling_Den_15_30_Tall = Seedling_15_30_Tall/0.0008,
    Seedling_Den_30_50_Tall = Seedling_30_50_Tall/0.0008,
    Seedling_Den_50_137_Tall = Seedling_50_137_Tall/0.0008,
    Sapling_Den_0_1_DBH = Sapling_0_1_DBH/0.008, 
    Sapling_Den_1_2half_DBH = Sapling_1_2half_DBH/0.008,
    Sapling_Den_2half_5_DBH = Sapling_2half_5_DBH/0.008,
    Sapling_Den_5_10_DBH = Sapling_5_10_DBH/0.008
  )

seed_sap_den <- seed_sap_den %>%
 dplyr::select(Plot_Code, Plant_Code, Seedling_Den_15_30_Tall, Seedling_Den_30_50_Tall, Seedling_Den_50_137_Tall, Sapling_Den_0_1_DBH, Sapling_Den_1_2half_DBH, Sapling_Den_2half_5_DBH, Sapling_Den_5_10_DBH)%>%
  replace(is.na(.), 0) %>%
  group_by(Plot_Code) %>%
  summarize(across(where(is.numeric), ~sum(.x, na.rm = TRUE))) %>%
  ungroup()
  

seed_sap_den <- seed_sap_den %>%
  group_by(Plot_Code) %>%
dplyr::mutate(seed_sap_Density = sum(Seedling_Den_15_30_Tall,
                                Seedling_Den_30_50_Tall,
                                Seedling_Den_50_137_Tall,Sapling_Den_0_1_DBH, 
                               Sapling_Den_1_2half_DBH,
                               Sapling_Den_2half_5_DBH, 
                               Sapling_Den_5_10_DBH)) %>%
  ungroup()

seed_sap_den <- seed_sap_den %>%
 dplyr::select(Plot_Code, seed_sap_Density)

seed_sap_top <- Reduce(function (...) { merge(..., all = TRUE) },  # Full join
                           list( seed_sap_other, seed_sap_top, seed_sap_den, cupn_carl))

```


#Mapping Regeneration by Composition
```{r}
baseman_carl = leaflet(data = seed_sap_top) %>%
  addProviderTiles("Esri.WorldImagery", options = providerTileOptions(noWrap = TRUE), group="Satellite") %>%
  addPolygons(data = imported_dat_carl, color = ~ "black") %>%
  addCircleMarkers(seed_sap_top$Longitude,
                   seed_sap_top$Latitude,  color = ~ "black", weight = 1, fillColor = ~"blue", fillOpacity = 1, radius = 4)

#SET VIEW SETS THE MIDDLE POINT OF THE MAP BUT DOES NOT NECESARRILY HELP FOR VIEWING PURPOSES
#  setView(lng = mean(seed_sap_top$Longitude), lat = mean(seed_sap_top$Latitude), zoom = 15)

baseman_carl <- baseman_carl %>%
  addMinicharts(seed_sap_top$Longitude, 
                seed_sap_top$Latitude, 
    chartdata =dplyr::select(seed_sap_top, Other:SASSALB), 
    type = "pie",
    width = sqrt(seed_sap_top$seed_sap_Density)/6.5,
    col = okabe_ito,
    showLabels = F) %>%
  addLabelOnlyMarkers(seed_sap_top$Longitude, 
                seed_sap_top$Latitude, label =  seed_sap_top$Plot_Code,
                      labelOptions = labelOptions(noHide = T, direction = 'top', textOnly = T, style=list('color'="#ffffff", 'fontSize'="10px"))) %>%
      fitBounds(min(seed_sap_top$Longitude), min(seed_sap_top$Latitude), max(seed_sap_top$Longitude), max(seed_sap_top$Latitude))


baseman_carl

#mapshot(baseman_carl, file = "Leaflet_CARLMap5.png")

```
#Creating Map 6, Tree Canopy Composition
```{r}
# Creating Tree (BA and Density columns by plot_code)
# Basal area in square meters = pi * (dbh^2) / 40000

#NEED TO BE LIVE, ANY TYPE OF CANOPY, CARL PLOTS, NO QAQC, CREATE BASAL AREA, Status code: 1 &3
```

```{r}
#tree_basics_count is the filtered dataset

tree_basics_count <- tree_basics %>%
dplyr::filter(Park_Code == "CARL")%>%
  dplyr::filter(Event_Type_Name != "QA/QC") %>%
dplyr::filter(Status_Code == 1) %>%
  dplyr::mutate(Basal_Area_ha = ((pi * DBH^2) /40000 )/0.04)

#Finding the top 11 most occuring species
top_tree <- tree_basics_count %>%
dplyr::mutate(Plant_Code = as.factor(Plant_Code)) %>%
  group_by(Plant_Code) %>%
  summarise(Observations = n()) %>%
  arrange(desc(Observations))  %>%
dplyr::mutate(Plant_Code = as.character(Plant_Code))

top_tree <- as.data.frame(top_tree[top_tree$Observations >20,])

listspecies <- as.list(top_tree[1])

#tree_count is the whole dataset subsetting to the top 10 species

tree_count <- tree_basics_count %>%
dplyr::filter(Plant_Code %in% listspecies$Plant_Code)

```

```{r}
tree_basics_other <- tree_basics_count %>%
dplyr::filter(!Plant_Code %in% listspecies$Plant_Code)

# tree_basics_other <- tree_basics_other %>%
#   replace(is.na(.), 0)

# tree_basics_other <- tree_basics_other %>%
# dplyr::mutate(Count = rowSums(across(Seedling_15_30_Tall:Sapling_5_10_DBH)))

tree_basics_other <- tree_basics_other %>%
  group_by(Plot_Code) %>%
  summarize(Other = n())
```


```{r}
#count number of observations of trees for tree density
tree_count <- tree_count %>%
  group_by(Plot_Code, Plant_Code) %>%
  summarize(DBH_Obs = n()) %>%
  ungroup()

tree_count <- tree_count %>%
  pivot_wider(names_from = Plant_Code, 
              values_from = DBH_Obs)

tree_ba<- tree_basics_count %>%
  group_by(Plot_Code) %>%
  summarise(Sum = sum(Basal_Area_ha))

tree_plot <-  Reduce(function (...) { merge(..., all = TRUE) },  # Full join
                           list(tree_ba, tree_basics_other, tree_count, cupn_carl))
```

```{r}
baseman_carl = leaflet(data = tree_plot) %>%
  addProviderTiles("Esri.WorldImagery", options = providerTileOptions(noWrap = TRUE), group="Satellite") %>%
  addPolygons(data = imported_dat_carl, color = ~ "black") %>%
  addCircleMarkers(tree_plot$Longitude,
                   tree_plot$Latitude,  color = ~ "black", weight = 1, fillColor = ~"blue", fillOpacity = 1, radius = 4)

#SET VIEW SETS THE MIDDLE POINT OF THE MAP BUT DOES NOT NECESARRILY HELP FOR VIEWING PURPOSES
#  setView(lng = mean(tree_plot$Longitude), lat = mean(tree_plot$Latitude), zoom = 15)

baseman_carl <- baseman_carl %>%
  addMinicharts(tree_plot$Longitude, 
                tree_plot$Latitude, 
    chartdata =dplyr::select(tree_plot, Other:LIRITUL), 
    type = "pie",
    width = sqrt(tree_plot$Sum) * 4,
    showLabels = F) %>%
  addLabelOnlyMarkers(tree_plot$Longitude, 
                tree_plot$Latitude, label =  tree_plot$Plot_Code,
                      labelOptions = labelOptions(noHide = T, direction = 'top', textOnly = T, style=list('color'="#ffffff", 'fontSize'="10px"))) %>%
      fitBounds(min(tree_plot$Longitude), min(tree_plot$Latitude), max(tree_plot$Longitude), max(tree_plot$Latitude))

baseman_carl
```
