---
title: "SFCN_Mapping"
output: html_document
date: '2022-12-05'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning = FALSE}
library(tidyr)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(plotly)
library(readr)
library(RColorBrewer)
library(knitr)
library(DT)
library(gt)
library(base)
library(stringr)
library(httr)
library(rgdal) 
library(sf) 
library(sp) 
library(rgeos)
library(leaflet)
library(leaflet.minicharts)
library(magrittr)
library(ggthemes)
library(gridExtra)

```

```{r}
setwd("C:/Users/kseeger/OneDrive - DOI/Desktop/GIT_RSTUDIO/SIP_NPS_CODE/Mapping/SFCN")

plot_locations <- read_csv("SFCNPlots_wLatLon_20221206.csv")
# - Plot locations with latitude and longitude

biscman_rich <- read_csv("SFCN_WoodyRich.csv")
# - BISCMAN live and dead DBH, and Densities, and proportion of species 
# Table 1

biscchh_rich <- read_csv("Book1.csv")
# - number of trees live and dead and their plot min & average DBH and density
#Table 2

#-----
#Table3 Split up

biscchh_species_total <- read_csv("SFCN_sp_total.csv")
# - Total count of unique species that are live and dead

biscchh_num_live_dead <- read_csv("SFCN_num_live_dead.csv")
# - BISCCHH count live and dead counts per species

biscchh_total_sp_rich <- read_csv("SFCN_Tot_sp_rich.csv")
# - total count of species per plot

```
# Manipulating original data frames
```{r}
biscchh_num_live_dead <- biscchh_num_live_dead %>% 
  replace(is.na(.), 0)
 
biscchh_num_live_dead

bisc_plot_location <- plot_locations %>%
  filter(stringr::str_starts(Plot_Code, "BISC"))

bisc_plot_location
```

#Downloading BISCMAN .shp shapefiles
```{r}
#useful website on how to create a CRS and sf/st/rgdal packages https://inbo.github.io/tutorials/tutorials/spatial_crs_coding/#:~:text=How%20to%20specify%20a%20CRS%20in%20R%3F%201,that%20might%20still%20work%2C%20their%20usage%20is%20discouraged%29.

# BISCMAN PLOTS
biscman_boundary <- st_read('./SFCN_Sapling_Frame_WGS8/Shapefiles_BISC_Forest_MangroveForest_SamplingFrame_GCS_WGS84.shp')

st_crs(biscman_boundary)

biscman_plot_locations <- bisc_plot_location %>%
  filter(grepl("BISCMAN", Plot_Code)) %>%
  select(Plot_Code, Latitude, Longitude)
```

#Downloading/Creating BISCCHH .shp shapefiles
```{r}
# BISCCHH PLOTS
biscchh_boundary <- st_read('./SFCN_Sapling_Frame_WGS8/Shapefiles_BISC_Forest_HardwoodHammocks_SamplingFrame_GCS_WGS84.shp')

st_crs(biscchh_boundary)

BISCCHH_plot_locations<- bisc_plot_location %>%
  filter(grepl("BISCCHH", Plot_Code)) %>%
  select(Plot_Code, Latitude, Longitude)

```
#Getting boundary for SFCN BISC Park
```{r}
park_code <- "BISC"
  # KATE, ENTER 4-LETTER PARK CODE HERE
  
unitBoundaryURL <- paste0("https://services1.arcgis.com/fBc8EJBxQRMcHlei/ArcGIS/rest/services/IMD_Units_Generic_areas_of_analysis_(AOAs)_-_IMD_BND_ALL_UNITS_AOA_nad_py_view/FeatureServer/0/query?where=UNITCODE+%3D+%27", park_code, "%27&objectIds=&time=&geometry=&geometryType=esriGeometryEnvelope&inSR=&spatialRel=esriSpatialRelIntersects&distance=&units=esriSRUnit_Meter&relationParam=&outFields=*&returnGeometry=true&maxAllowableOffset=&geometryPrecision=&outSR=4326&gdbVersion=&returnDistinctValues=false&returnIdsOnly=false&returnCountOnly=false&returnExtentOnly=false&orderByFields=&groupByFieldsForStatistics=&outStatistics=&returnZ=false&returnM=false&resultOffset=&resultRecordCount=&f=geojson") # save output as WGS84

paste(unitBoundaryURL)
```
#Creating geojson file 
```{r}
tempUnitOutput <- "tempunit.geojson"

#checking error code below
bisc_plot<- tryCatch(readOGR(dsn = tempUnitOutput, dropNULLGeometries = FALSE), error=function(e) print("Error retrieving data")) # return error message if problems still arise with downloading from web services

#shiny::req(class(imported_dat)=="SpatialPolygonsDataFrame")
bisc_plot<- spTransform(bisc_plot, CRS("+proj=longlat +datum=WGS84")) # convert to WGS84

bisc_plot<- readOGR(dsn = tempUnitOutput, dropNULLGeometries = FALSE) 

#plot(BISC_PLOT)
```

#Compiling and mutating data for the final maps
```{r}
#PLOTCODE_RICH = lat and long with min, max, live , dead, and dbh data. 
#PLOTCODE_NUM_Trees are datasets ready to be used with leaflet and include lat, long, and number of live and dead trees
```

#BISCMAN Live/Dead Prop
```{r}
# Creating biscman_rich which appends lat and long to data set
biscman_rich <- biscman_rich %>%
  rename(Plot_Code = Site)

biscman_rich <- merge(x = biscman_rich, y = biscman_plot_locations[ , c("Plot_Code", "Latitude", "Longitude")], by ='Plot_Code', all.x=TRUE)

biscman_num_trees <- biscman_rich %>%
  dplyr::select(Plot_Code, Status, Num_Trees, Latitude, Longitude) %>%
  dplyr::mutate(Latitude = as.double(Latitude), 
                Longitude = as.double(Longitude))

#Pivoting table
biscman_num_trees <- biscman_num_trees %>%
  pivot_wider(names_from = Status,
              values_from = Num_Trees)
```

#BISCCHH Live/Dead Trees
```{r}
#BISCCHH NUM TREES 
biscchh_num_trees <- biscchh_rich %>%
  rename(Plot_Code = Site) %>%
  dplyr::select(Plot_Code, status, Num_Trees)

BISCCHH_plot_locations <- plot_locations %>%
  dplyr::filter(stringr::str_starts(Plot_Code, "BISCCHH")) %>%
  dplyr::select(Plot_Code, Latitude, Longitude)


#Merging data with lat and long to plot it
biscchh_num_trees <- left_join(biscchh_num_trees, BISCCHH_plot_locations, by = "Plot_Code")


biscchh_num_trees <- biscchh_num_trees %>%
  pivot_wider(names_from = status, 
              values_from = Num_Trees 
              )

biscman_num_trees
biscchh_num_trees

```

#creating official leaflet map showing the two forest communities in BISC and showing proportional data from both BISCCHH and BISCMAN
```{r}
#Brown and green colors
colors = c("#008000", "#8d5542")

BISC_MAP = leaflet(data = bisc_plot_location) %>%
  addTiles() %>%
#  addPolygons(data = BISC_PLOT, color = ~ "black", weight = 1.5, fillOpacity = 0) %>%
#  setView(lng = mean(biscman_num_trees$Longitude), lat = mean(biscman_num_trees$Latitude), zoom = 11.35)%>%
  addPolygons(data = biscchh_boundary, color = ~"black", weight = 1, fillColor = ~"yellow", fillOpacity = 0.6) %>%
  addPolygons(data = biscman_boundary, color = ~ "black", weight = 1, fillColor = ~"red", fillOpacity = 0.6)

BISC_MAP %>%
  addMinicharts(biscchh_num_trees$Longitude, 
                biscchh_num_trees$Latitude, 
                chartdata = dplyr::select(biscchh_num_trees, Live, Dead), 
                colorPalette = colors, 
                type = "pie", 
                width = sqrt(biscchh_num_trees$Live + biscchh_num_trees$Dead) * 3) %>%
  addMinicharts(biscman_num_trees$Longitude, 
                biscman_num_trees$Latitude, 
                chartdata = dplyr::select(biscman_num_trees, Live, Dead), 
                colorPalette = colors, 
                type = "pie", 
                width = sqrt(biscman_num_trees$Live + biscman_num_trees$Dead) * 3) %>%
  addLabelOnlyMarkers(biscchh_num_trees$Longitude, 
                biscchh_num_trees$Latitude, label =  biscchh_num_trees$Plot_Code,
                      labelOptions = labelOptions(noHide = T, direction = 'top', textOnly = T, style=list('color'="#171717", 'fontSize'="10px"))) %>%
   addLabelOnlyMarkers(biscman_num_trees$Longitude, 
                biscman_num_trees$Latitude, label =  biscman_num_trees$Plot_Code,
                      labelOptions = labelOptions(noHide = T, direction = 'top', textOnly = T, style=list('color'="#171717", 'fontSize'="10px"))) %>%
    fitBounds(min(biscman_num_trees $Longitude), 
            min(biscman_num_trees $Latitude), 
            max(biscchh_num_trees $Longitude),
            max(biscman_num_trees $Latitude))

  
```

```{r}
#Brown and green colors
colors = c("#008000", "#8d5542")

BISC_MAP = leaflet(data = bisc_plot_location) %>%
  addTiles() %>%
  setView(lng = mean(biscman_num_trees$Longitude), lat = mean(biscman_num_trees$Latitude), zoom = 11)%>%
  addPolygons(data = bisc_plot, color = ~ "black", weight = 1.5, fillOpacity = 0) %>%
  addPolygons(data = biscchh_boundary, color = ~"black", weight = 1, fillColor = ~"yellow", fillOpacity = 0.6) %>%
  addPolygons(data = biscman_boundary, color = ~ "black", weight = 1, fillColor = ~"red", fillOpacity = 0.6)

BISC_MAP %>%
  addMinicharts(biscchh_num_trees$Longitude, 
                biscchh_num_trees$Latitude, 
                chartdata = dplyr::select(biscchh_num_trees, Live, Dead), 
                colorPalette = colors, 
                type = "pie", 
                width = sqrt(biscchh_num_trees$Live + biscchh_num_trees$Dead) * 3) %>%
  addMinicharts(biscman_num_trees$Longitude, 
                biscman_num_trees$Latitude, 
                chartdata = dplyr::select(biscman_num_trees, Live, Dead), 
                colorPalette = colors, 
                type = "pie", 
                width = sqrt(biscman_num_trees$Live + biscman_num_trees$Dead) * 3) %>%
  addLabelOnlyMarkers(biscchh_num_trees$Longitude, 
                biscchh_num_trees$Latitude, label =  biscchh_num_trees$Plot_Code,
                      labelOptions = labelOptions(noHide = T, direction = 'top', textOnly = T, style=list('color'="#171717", 'fontSize'="10px"))) %>%
   addLabelOnlyMarkers(biscman_num_trees$Longitude, 
                biscman_num_trees$Latitude, label =  biscman_num_trees$Plot_Code,
                      labelOptions = labelOptions(noHide = T, direction = 'top', textOnly = T, style=list('color'="#171717", 'fontSize'="10px")))

  
```

#BISCMAN
```{r}
#Brown and green colors
colors = c("#008000", "#8d5542")

BISC_MAP = leaflet(data = bisc_plot_location) %>%
  addProviderTiles("Esri.WorldImagery", options = providerTileOptions(noWrap = TRUE), group="Satellite") %>%  
  setView(lng = mean(biscman_num_trees$Longitude), lat = mean(biscman_num_trees$Latitude), zoom = 11)%>%
 # addPolygons(data = BISC_PLOT, color = ~ "black", weight = 1.5, fillOpacity = 0) %>%
  #addPolygons(data = biscchh_boundary, color = ~"black", weight = 1, fillColor = ~"yellow", fillOpacity = 0.6) %>%
  addPolygons(data = biscman_boundary, color = ~ "black", weight = 1, fillColor = ~"red", fillOpacity = 0.6)

BISC_MAP %>%
  addMinicharts(biscman_num_trees$Longitude, 
                biscman_num_trees$Latitude, 
                chartdata = dplyr::select(biscman_num_trees, Live, Dead), 
                colorPalette = colors, 
                type = "pie", 
                width = sqrt(biscman_num_trees$Live + biscman_num_trees$Dead) * 3) %>%
   addLabelOnlyMarkers(biscman_num_trees$Longitude, 
                biscman_num_trees$Latitude, label =  biscman_num_trees$Plot_Code,
                      labelOptions = labelOptions(noHide = T, direction = 'top', textOnly = T, style=list('color'="#ffffff", 'fontSize'="10px")))
```

#BISCCHH
```{r}
#Brown and green colors
colors = c("#008000", "#8d5542")

BISC_MAP = leaflet(data = bisc_plot_location) %>%
  addProviderTiles(providers$Esri.WorldImagery) %>%
  setView(lng = mean(biscchh_num_trees$Longitude), lat = mean(biscchh_num_trees$Latitude), zoom = 11)%>%
#  addPolygons(data = BISC_PLOT, color = ~ "black", weight = 1.5, fillOpacity = 0) %>%
  addPolygons(data = biscchh_boundary, color = ~"black", weight = 1, fillColor = ~"yellow", fillOpacity = 0.6) #%>%
 # addPolygons(data = biscman_boundary, color = ~ "black", weight = 1, fillColor = ~"red", fillOpacity = 0.6)

BISC_MAP %>%
  addMinicharts(biscchh_num_trees$Longitude, 
                biscchh_num_trees$Latitude, 
                chartdata = dplyr::select(biscchh_num_trees, Live, Dead), 
                colorPalette = colors, 
                type = "pie", 
                width = sqrt(biscchh_num_trees$Live + biscchh_num_trees$Dead) * 3) %>%
  addLabelOnlyMarkers(biscchh_num_trees$Longitude, 
                biscchh_num_trees$Latitude, label =  biscchh_num_trees$Plot_Code,
                      labelOptions = labelOptions(noHide = T, direction = 'top', textOnly = T, style=list('color'="#ffffff", 'fontSize'="10px", "text-shadow" = "#ffffff'"
          )
        )
      )
```

#BISCMAN
```{r}
six <- c("#56B4E9",  "#0072B2", "#E69F00", "#D55E00", '#FFAABB', '#882255')

biscman_num_sp <- biscman_rich %>%
  dplyr::select(Plot_Code, Status, Latitude, Longitude, `Avicennia germinans`, `Laguncularia racemosa`, `Rhizophora mangle`)

biscman_num_sp <- biscman_num_sp %>%
  pivot_wider(names_from = Status, 
              values_from= c(`Avicennia germinans`, `Laguncularia racemosa`, `Rhizophora mangle`))

BISC_MAP = leaflet(data = bisc_plot_location) %>%
  addProviderTiles("Esri.WorldImagery", options = providerTileOptions(noWrap = TRUE), group="Satellite") %>%  
  setView(lng = mean(biscman_num_sp$Longitude), lat = mean(biscman_num_sp$Latitude), zoom = 11)%>%
  addPolygons(data = biscman_boundary, color = ~ "black", weight = 1, fillColor = ~"red", fillOpacity = 0.6)

BISC_MAP %>%
  addMinicharts(biscman_num_sp$Longitude, 
                biscman_num_sp$Latitude, 
                chartdata = dplyr::select(biscman_num_sp, `Avicennia germinans_Live`, `Avicennia germinans_Dead`, `Laguncularia racemosa_Live`, `Laguncularia racemosa_Dead`, `Rhizophora mangle_Live`, `Rhizophora mangle_Dead`), 
                colorPalette= six,
                height = 45, width = 60)%>%
   addLabelOnlyMarkers(biscman_num_sp$Longitude, 
                biscman_num_sp$Latitude, label =  biscman_num_sp$Plot_Code,
                      labelOptions = labelOptions(noHide = T, direction = 'top', textOnly = T, style=list('color'="#ffffff", 'fontSize'="10px")))

```

